### 2017-07-03 WekaTesting/stats/Time_Fig.tex
Updated post hoc script
> added option for setting time as min for best value
> ran and updated tex files

\begin{tikzpicture}
\draw (1,0) -- (12,0);
\foreach \x in {1,2,3,4,5,6,7,8,9,10,11,12} {
\draw (\x, 0) -- ++(0,.1) node [above,scale=0.7] {\tiny \x};
\ifthenelse{\x < 12}{\draw (\x+.5, 0) -- ++(0,.03);}{}
}

\coordinate (c0) at (2.2667,0);
\coordinate (c1) at (6.2,0);
\coordinate (c2) at (3.1,0);
\coordinate (c3) at (7.2,0);
\coordinate (c4) at (11.0667,0);
\coordinate (c5) at (4.3,0);
\coordinate (c6) at (8.5333,0);
\coordinate (c7) at (10.1333,0);
\coordinate (c8) at (1.8667,0);
\coordinate (c9) at (7.2,0);
\coordinate (c10) at (6.4667,0);
\coordinate (c11) at (9.6667,0);

\node (l0) at (c0) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIRSVM};
\node (l1) at (c1) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny miGraph};
\node (l2) at (c2) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIBoost};
\node (l3) at (c3) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIOptimalBall};
\node (l4) at (c4) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIDD};
\node (l5) at (c5) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIWrapper};
\node (l6) at (c6) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MISMO};
\node (l7) at (c7) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MISVM};
\node (l8) at (c8) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny SimpleMI};
\node (l9) at (c9) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny TLC};
\node (l10) at (c10) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny Bagging};
\node (l11) at (c11) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny Stacking};

\fill[fill=gray,fill opacity=0.5] (1.8667,-0.1) rectangle (2.8667,0.1);

\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12} {
\draw (l\x) -| (c\x);
};
\end{tikzpicture}