### 2017-07-03 WekaTesting/stats/metaranks_Fig.tex

\begin{tikzpicture}
\draw (1,0) -- (10,0);
\foreach \x in {1,2,3,4,5,6,7,8,9,10} {
\draw (\x, 0) -- ++(0,.1) node [above,scale=0.7] {\tiny \x};
\ifthenelse{\x < 10}{\draw (\x+.5, 0) -- ++(0,.03);}{}
}

\coordinate (c0) at (2.1667,0);
\coordinate (c1) at (3.3333,0);
\coordinate (c2) at (8.9167,0);
\coordinate (c3) at (7.75,0);
\coordinate (c4) at (9.25,0);
\coordinate (c5) at (9.0833,0);
\coordinate (c6) at (5.5,0);
\coordinate (c7) at (8.8333,0);
\coordinate (c8) at (7.3333,0);
\coordinate (c9) at (5.0833,0);
\coordinate (c10) at (4.0833,0);
\coordinate (c11) at (6.6667,0);

\node (l0) at (c0) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIRSVM};
\node (l1) at (c1) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny miGraph};
\node (l2) at (c2) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIBoost};
\node (l3) at (c3) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIOptimalBall};
\node (l4) at (c4) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIDD};
\node (l5) at (c5) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MIWrapper};
\node (l6) at (c6) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MISMO};
\node (l7) at (c7) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny MISVM};
\node (l8) at (c8) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny SimpleMI};
\node (l9) at (c9) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny TLC};
\node (l10) at (c10) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny Bagging};
\node (l11) at (c11) [above left=.1cm and 0cm, align=right,scale=0.7] {\tiny Stacking};

\fill[fill=gray,fill opacity=0.5] (2.1667,-0.1) rectangle (3.1667,0.1);

\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12} {
\draw (l\x) -| (c\x);
};
\end{tikzpicture}